---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# ovbias

<!-- badges: start -->
<!-- badges: end -->

The goal of ovbias is to compute quantiles of the empirical distribution of the bias-adjusted treatment effect in a linear econometric model with omitted variable bias. To analyze such models, one can consider three regression models: (a) a short regression model where the outcome variable is regressed on the treatment variable, with or without additional controls; (b) an intermediate regression model where additional control variables are added to the short regression; (c) a hypothetical long regression model, where and index of the omitted variable(s) is added to the intermediate regressions. The selection on unobservables is captured by the parameter `delta`; and the R-squared in the hypothetical long regression is captured by the parameter `Rmax`.

The main function in this package runs three regressions: (a) the short regression, (b) the intermediate regression, and (c) an auxiliary regression where the treatment variable is regression on all included control variables. Relevant parameters are collected from these three regressions and used to construct a cubic equation in the omitted variable bias. The cubic equation is solved on a NxN grid over a bounded box of the (`delta`,`Rmax`) plane, and the roots of the cubic are used to construct the bias-adjusted treatment effect. 

The bounded box of the (`delta`,`Rmax`) plane is defined by the Cartesian products of two intervals: [`deltalow`, `deltahigh`], and [`Rtilde`, `Rmax`]. We posit two possibilities for `Rmax`: (a) `1.3*Rtilde`, (b) `2.47*Rtilde`. Hence, we can define 4 different regimes: (1) low `delta`, low `Rmax`, where `deltalow < delta < 1` and `Rtilde < Rmax < 1.30 Rtilde`; (2) low `delta`, high `Rmax`, where `deltalow < delta < 1` and `Rtilde < Rmax < 2.47 Rtilde`; (3) high `delta`, low `Rmax`, where `1 < delta < deltahigh` and `Rtilde < Rmax < 1.30 Rtilde`; (4) high `delta`, high `Rmax`, where `1 < delta < deltahigh` and `Rtilde < Rmax < 2.47 Rtilde`.

The output of the function is a list of two elements. The first element is a data frame of the parameters extracted from the three regressions; the second element of the list is a matrix containing the empirical distribution of the bias-adjusted treatment effect for the four regimes. 

For more details see Basu, D. (2021). "Bounding Sets for Treatment Effects with Proportional Selection". Economics Department Working Paper Series. 307. University of Massachusetts Amhers. URL: https://scholarworks.umass.edu/econ_workingpaper/307  

## Installation

You can install ovbias like so:

``` r
# install.packages("devtools")
devtools::install_github("dbasu-umass/ovbias")
```
## Example

Let us create a simulated data set.

```{r example}
library(ovbias)

# Parameters
a0 <- 0.5
a1 <- 0.25
a2 <- 0.75
a3 <- 1.0
a4 <- -0.75 
a5 <- -0.25

# Regressors
x1 <- rnorm(100, mean=0, sd=1)
x2 <- rnorm(100, mean=1, sd=2)
x3 <- rnorm(100, mean=2, sd=3)
x4 <- rnorm(100, mean=3, sd=4)
x5 <- rep(c(0,1),50)

# Outcome variable
y <- a0 + a1*x1 + a2*x2 + a3*x3 + a4*x4 + a5*x5

# Create data frame
d1 <- data.frame(y=y,x1=x1,x2=x2,x3=x3,x4=x4,x5=x5)
```

The researcher is interested in estimating the effect of the treatment variable, `x1`, on the outcome variable, `y`. Suppose the researcher is unable to collect data on the following two variables `x4` and `x5`. When the researcher estimates a linear model, the treatment effect is estimated with bias (because of the two omitted variables). 

We would like to use the `bset2()` function to find bounding sets of the bias-adjusted treatment effect. We will use `deltalow=0.01` and `deltahigh=5.0` in our call of the `bset2()` function.

```{r}
# Generate results
bset2(
  data = d1,
  outcome = "y",
  treatment = "x1",
  shortreg = y ~ x1,
  intreg = y ~ x1 + x2 + x3,
  auxreg = x1 ~ x2 + x3, 
  deltalow = 0.01, 
  deltahigh=5.00
  )

```

To see the relevant parameters that emerged from the short, intermediate and auxiliary regressions, look at the first element of the list, `$bsetpar`; to see the quantiles of the empirical distribution of the bias-adjusted treatment effect look at the second element of the list, `$bsetqnt`. 
